CF Registry Analysis

### 2022-12-27 Christine Chan/CF Registry/cf_registry_analysis.qmd
updates finished!

```{r}
# All data
encounter <- read.csv("Z:/PEDS/RI Biostatistics Core/Shared/Shared Projects/Laura/Peds Endo/Christine Chan/CF registry data/CFF19_encountersMerged_Del1.csv",
                      na.strings = c("", "N/A")
)
annual <- read.csv("Z:/PEDS/RI Biostatistics Core/Shared/Shared Projects/Laura/Peds Endo/Christine Chan/CF registry data/CFF19_Annualized_Del1.csv",
                   na.strings = c("", "N/A")
)
encounter <- encounter %>%
  arrange(eDWID, reviewyear, encounternum)
annual = annual %>% arrange(eDWID,ReviewYear)
# Add age
ages <- annual %>% 
  rename(reviewyear = ReviewYear) %>%
  group_by(eDWID,reviewyear) %>%
  summarise(Age = as.numeric(Age_YrEnd[1]),.groups = "drop")
encounter = left_join(encounter,ages,by = c("eDWID", "reviewyear"))
encounter$`Age Group` = cut(encounter$Age,
                            c(-Inf,10,19,31,Inf),
                            labels = c("< 10", "10 - 18", "19 - 30", "> 30"),
                            right=F)
# For those with CFRD by OGTT, find first visit with diagnosis
first_cfrd_ogtt <- encounter %>% 
  filter(twohour_bloodglucose >= 200) %>%
  group_by(eDWID) %>%
  filter(row_number()==1) %>%
  select(eDWID, reviewyear, `Age Group`) %>%
  rename(first_cfrd_ogtt = reviewyear)
# For those with CFRD by registry criteria, find first visit with diagnosis
first_cfrd_cffpr <- encounter %>% 
  filter(cfrd_status==3) %>%
  group_by(eDWID) %>%
  filter(row_number()==1) %>%
  select(eDWID, reviewyear, `Age Group`) %>%
  rename(first_cfrd_cffpr = reviewyear)
# Both
first_cfrd_both <- encounter %>% 
  filter(cfrd_status==3|twohour_bloodglucose >= 200) %>%
  group_by(eDWID) %>%
  filter(row_number()==1) %>%
  select(eDWID, reviewyear, `Age Group`) %>%
  rename(first_cfrd_both = reviewyear)
# Add to encounter level data
encounter <- left_join(encounter, first_cfrd_ogtt, by = c("eDWID", "Age Group"))
encounter <- left_join(encounter, first_cfrd_cffpr, by = c("eDWID", "Age Group"))
encounter <- left_join(encounter, first_cfrd_both,by = c("eDWID", "Age Group"))
# Fill down CFRD Yes/No by year
encounter <- encounter %>%
  group_by(eDWID) %>%
  mutate(
    cfrd_ogtt = case_when(
      reviewyear < first_cfrd_ogtt ~ "No",
      reviewyear >= first_cfrd_ogtt ~ "Yes"
    ),
    cfrd_cffpr = case_when(
      reviewyear < first_cfrd_cffpr ~ "No",
      reviewyear >= first_cfrd_cffpr ~ "Yes"
    ),
    cfrd_both = case_when(
      reviewyear < first_cfrd_both ~ "No",
      reviewyear >= first_cfrd_both ~ "Yes"
    )
  ) %>%
  ungroup()
encounter$cfrd_ogtt[is.na(encounter$cfrd_ogtt)] <- "No"
encounter$cfrd_cffpr[is.na(encounter$cfrd_cffpr)] <- "No"
encounter$cfrd_both[is.na(encounter$cfrd_both)] <- "No"
# Ordered factors
encounter$cfrd_ogtt = factor(encounter$cfrd_ogtt,levels = c("No","Yes"),ordered = T)
encounter$cfrd_cffpr = factor(encounter$cfrd_cffpr,levels = c("No","Yes"),ordered = T)
encounter$cfrd_both = factor(encounter$cfrd_both,levels = c("No","Yes"),ordered = T)
# Remove those missing age
encounter = encounter %>% filter(!is.na(Age))
```

```{r}
#| label: tbl-incidence_cffpr
#| tbl-cap: "Year-on-Year Incidence of CFRD Based on CFFPR Alone"
# For each year, pull everyone with and without CFRD that year
year_inc = function(df=encounter,year,py = 1000){
  # Filter and count
  inc = df %>%  
    filter(reviewyear==year,is.na(first_cfrd_cffpr) | first_cfrd_cffpr == year) %>%
    group_by(eDWID) %>% filter(row_number() == 1) %>% 
    select(eDWID,`Age Group`,cfrd_cffpr)
  inc = inc %>% group_by(cfrd_cffpr) %>% count(`Age Group`,.drop = F)
  # Calculations
  new_cases = sum(inc$n[inc$cfrd_cffpr=="Yes"])
  total_at_risk = sum(inc$n)
  total_inc = new_cases/(total_at_risk/py)

@@

       `Total New Cases Age in 2018 < 10` = new_cases_10,
       `Total At Risk Age in 2018 < 10` = at_risk_10,
       `Total Incidence Age in 2018 < 10` = inc_10,

       `Total New Cases Age in 2018 10-18` = new_cases_10_18,
       `Total At Risk Age in 2018 10-18` = at_risk_10_18,
       `Total Incidence Age in 2018 10-18` = inc_10_18,

       `Total New Cases Age in 2018 19-30` = new_cases_19_30,
       `Total At Risk Age in 2018 19-30` = at_risk_19_30,
       `Total Incidence Age in 2018 19-30` = inc_19_30,

       `Total New Cases Age in 2018 > 30` = new_cases_30,
       `Total At Risk Age in 2018 > 30` = at_risk_30,
       `Total Incidence Age in 2018 > 30` = inc_30)
}
incidence = lapply(2003:2018, function(y){year_inc(year = y)})
incidence = data.frame(do.call(rbind,incidence),check.names = F)
incidence = data.frame(lapply(incidence,as.numeric),check.names = F)
kable(incidence,digits = 3)
```

@@

```{r}
#| label: tbl-prevalence_cffpr
#| tbl-cap: "Yearly Prevalence of CFRD Based on CFFPR Alone"
encounter %>% group_by(eDWID,reviewyear) %>% 
  summarise(cfrd = max(cfrd_cffpr),.groups = "drop") %>%
  group_by(reviewyear) %>%
  summarise(Prevalence = mean(as.numeric(cfrd)-1,na.rm = T)*100) %>%
  kable(.,digits = 3)
```

In @tbl-prevalence_cffpr, prevalence is calculated as the total number of CFRD cases (by registry definition alone) seen in clinic divided by the total number of people seen in clinic in a given year.

```{r}
#| label: tbl-incidence_both
#| tbl-cap: "Year-on-Year Incidence of CFRD Based on CFFPR and OGTT"
# For each year, pull everyone with and without CFRD that year
year_inc = function(df=encounter,year,py = 1000){
  # Filter and count
  inc = df %>% 
    filter(reviewyear==year,is.na(first_cfrd_both) | first_cfrd_both == year) %>%
    group_by(eDWID) %>% filter(row_number() == 1) %>%
    select(eDWID,reviewyear,encounternum,`Age Group`,
           twohour_bloodglucose,cfrd_both,first_cfrd_both)
  inc = inc %>% group_by(cfrd_both) %>% count(`Age Group`,.drop = F)
  # Calculations
  new_cases = sum(inc$n[inc$cfrd_both=="Yes"])
  total_at_risk = sum(inc$n)
  total_inc = new_cases/(total_at_risk/py)

@@

```{r}
#| label: tbl-prevalence_both
#| tbl-cap: "Yearly Prevalence of CFRD Based on CFFPR and OGTT"
encounter$cfrd_both = factor(encounter$cfrd_both,levels = c("No","Yes"),ordered = T)
encounter %>% group_by(eDWID,reviewyear) %>% 
  summarise(cfrd = max(cfrd_both),.groups = "drop") %>%
  group_by(reviewyear) %>%
  summarise(`Prevalence` = mean(as.numeric(cfrd)-1,na.rm = T)*100) %>%
  kable(.,digits = 3)
```

In @tbl-prevalence_both, prevalence is calculated as the total number of CFRD cases (by either registry definition or OGTT) seen in clinic divided by the total number of people seen in clinic in a given year.

```{r}
#| label: tbl-screening
#| tbl-cap: "Yearly Screening Rates"
year_screen = function(df=encounter,year){
  # Filter and count
  inc = df %>% 
    filter(reviewyear==year,first_cfrd_both >= year) %>%
    mutate(ogtt = !(is.na(ogttfast_bloodglucose) | is.na(twohour_bloodglucose))) %>%
    group_by(eDWID) %>%
    summarise(ogtt = max(ogtt),`Age Group` = `Age Group`[1])
  inc = inc %>% group_by(ogtt) %>% count(`Age Group`) %>% filter(`Age Group` != "< 10")
  # Calculations
  total_screened = sum(inc$n[inc$ogtt==1])
  total_seen = sum(inc$n)
  screen_rate = total_screened/total_seen * 100

@@

In @tbl-screening, people with CFRD (by either registry definition or OGTT) and those under 10 were excluded. Screening rate is calculated as the number of people with an OGTT divided by the total number of people seen in clinic that year.

```{r}
#| label: tbl-screening_a1c
#| tbl-cap: "Yearly Screening Rates"
year_screen = function(df=encounter,year){
  # Filter and count
  inc = df %>% 
    filter(reviewyear==year,first_cfrd_both >= year) %>%
    mutate(hba1c = !is.na(hgba1c)) %>%
    group_by(eDWID) %>%
    summarise(hba1c = max(hba1c),`Age Group` = `Age Group`[1])
  inc = inc %>% group_by(hba1c) %>% count(`Age Group`) %>% filter(`Age Group` != "< 10")
  # Calculations
  total_screened = sum(inc$n[inc$hba1c==1])
  total_seen = sum(inc$n)
  screen_rate = total_screened/total_seen * 100

  screened_10_18 = inc$n[inc$hba1c==1 & inc$`Age Group` == "10 - 18"]
  seen_10_18 = sum(inc$n[inc$`Age Group` == "10 - 18"])
  rate_10_18 = screened_10_18/seen_10_18 *100

  screened_19_30 = inc$n[inc$hba1c==1 & inc$`Age Group` == "19 - 30"]
  seen_19_30 = sum(inc$n[inc$`Age Group` == "19 - 30"])
  rate_19_30 = screened_19_30/seen_19_30 *100

  screened_30 = inc$n[inc$hba1c==1 & inc$`Age Group` == "> 30"]
  seen_30 = sum(inc$n[inc$`Age Group` == "> 30"])
  rate_30 = screened_30/seen_30 *100

@@

In @tbl-screening_a1c, people with CFRD (by either registry definition or OGTT) and those under 10 were excluded. Screening rate is calculated as the number of people with an HbA1c value divided by the total number of people seen in clinic that year.

>>>>

```{r}
# All data
encounter <- read.csv("Z:/PEDS/RI Biostatistics Core/Shared/Shared Projects/Laura/Peds Endo/Christine Chan/CF registry data/CFF19_encountersMerged_Del1.csv",
                      na.strings = c("", "N/A")
)
annual <- read.csv("Z:/PEDS/RI Biostatistics Core/Shared/Shared Projects/Laura/Peds Endo/Christine Chan/CF registry data/CFF19_Annualized_Del1.csv",
                   na.strings = c("", "N/A")
)
encounter <- encounter %>%
  arrange(eDWID, reviewyear, encounternum)
annual = annual %>% arrange(eDWID,ReviewYear)
# Add age
ages <- annual %>% 
  mutate(`Age in 2018` = 2018-(ReviewYear-Age_YrEnd)) %>%
  rename(reviewyear = ReviewYear) %>%
  group_by(eDWID,reviewyear) %>%
  summarise(`Age in 2018` = as.numeric(names(sort(table(`Age in 2018`), 
                                                  decreasing = TRUE)[1])),
            Age = Age_YrEnd[1],.groups = "drop")
encounter = left_join(encounter,ages,by = c("eDWID", "reviewyear"))
encounter$`Age Group 2018` = cut(encounter$`Age in 2018`,
                            c(-Inf,10,19,31,Inf),
                            labels = c("< 10", "10 - 18", "19 - 30", "> 30"),
                            right=F)
encounter$`Age Group` = cut(encounter$Age,
                            c(-Inf,10,19,31,Inf),
                            labels = c("< 10", "10 - 18", "19 - 30", "> 30"),
                            right=F)
# For those with CFRD by OGTT, find first visit with diagnosis
first_cfrd_ogtt <- encounter %>% 
  filter(twohour_bloodglucose >= 200) %>%
  group_by(eDWID) %>%
  filter(row_number()==1) %>%
  select(eDWID, reviewyear, `Age in 2018`) %>%
  rename(first_cfrd_ogtt = reviewyear)
# For those with CFRD by registry criteria, find first visit with diagnosis
first_cfrd_cffpr <- encounter %>% 
  filter(cfrd_status==3) %>%
  group_by(eDWID) %>%
  filter(row_number()==1) %>%
  select(eDWID, reviewyear, `Age in 2018`) %>%
  rename(first_cfrd_cffpr = reviewyear)
# Both
first_cfrd_both <- encounter %>% 
  filter(cfrd_status==3|twohour_bloodglucose >= 200) %>%
  group_by(eDWID) %>%
  filter(row_number()==1) %>%
  select(eDWID, reviewyear, `Age in 2018`) %>%
  rename(first_cfrd_both = reviewyear)
# Add to encounter level data
encounter <- left_join(encounter, first_cfrd_ogtt, by = c("eDWID", "Age in 2018"))
encounter <- left_join(encounter, first_cfrd_cffpr, by = c("eDWID", "Age in 2018"))
encounter <- left_join(encounter, first_cfrd_both,by = c("eDWID", "Age in 2018"))
# Fill down CFRD Yes/No by year
encounter <- encounter %>%
  group_by(eDWID) %>%
  mutate(
    cfrd_ogtt = case_when(
      reviewyear < first_cfrd_ogtt ~ "No",
      reviewyear >= first_cfrd_ogtt ~ "Yes"
    ),
    cfrd_cffpr = case_when(
      reviewyear < first_cfrd_cffpr ~ "No",
      reviewyear >= first_cfrd_cffpr ~ "Yes"
    ),
    cfrd_both = case_when(
      reviewyear < first_cfrd_both ~ "No",
      reviewyear >= first_cfrd_both ~ "Yes"
    )
  ) %>%
  ungroup()
encounter$cfrd_ogtt[is.na(encounter$cfrd_ogtt)] <- "No"
encounter$cfrd_cffpr[is.na(encounter$cfrd_cffpr)] <- "No"
encounter$cfrd_both[is.na(encounter$cfrd_both)] <- "No"
# Ordered factors
encounter$cfrd_ogtt = factor(encounter$cfrd_ogtt,levels = c("No","Yes"),ordered = T)
encounter$cfrd_cffpr = factor(encounter$cfrd_cffpr,levels = c("No","Yes"),ordered = T)
encounter$cfrd_both = factor(encounter$cfrd_both,levels = c("No","Yes"),ordered = T)
# Remove those missing age
encounter = encounter %>% filter(!is.na(Age))
```

```{r}
#| label: tbl-incidence_cffpr
#| tbl-cap: "Year-on-Year Incidence of CFRD Based on CFFPR Alone"
# For each year, pull everyone with and without CFRD that year
year_inc = function(df=encounter,year,py = 1000){
  # Filter and count
  inc = df %>%  
    filter(reviewyear==year,is.na(first_cfrd_cffpr) | first_cfrd_cffpr == year) %>%
    group_by(eDWID) %>% filter(row_number() == 1) %>% 
    select(eDWID,`Age Group`,cfrd_cffpr)
  inc = inc %>% group_by(cfrd_cffpr) %>% count(`Age Group`,.drop = F)
  inc = inc[!is.na(inc$`Age Group`),]
  # Calculations
  new_cases = sum(inc$n[inc$cfrd_cffpr=="Yes"])
  total_at_risk = sum(inc$n)
  total_inc = new_cases/(total_at_risk/py)

@@

       `Total New Cases Age < 10` = new_cases_10,
       `Total At Risk Age < 10` = at_risk_10,
       `Total Incidence Age < 10` = inc_10,

       `Total New Cases Age 10-18` = new_cases_10_18,
       `Total At Risk Age 10-18` = at_risk_10_18,
       `Total Incidence Age 10-18` = inc_10_18,

       `Total New Cases Age 19-30` = new_cases_19_30,
       `Total At Risk Age 19-30` = at_risk_19_30,
       `Total Incidence Age 19-30` = inc_19_30,

       `Total New Cases Age > 30` = new_cases_30,
       `Total At Risk Age > 30` = at_risk_30,
       `Total Incidence Age > 30` = inc_30)
}
incidence = lapply(2003:2018, function(y){year_inc(year = y)})
incidence = data.frame(do.call(rbind,incidence),check.names = F)
incidence = data.frame(lapply(incidence,as.numeric),check.names = F)
kable(incidence,digits = 3)
```

@@

```{r}
#| label: tbl-prevalence_cffpr_year
#| tbl-cap: "Yearly Prevalence of CFRD by Age Group Based on CFFPR Alone"
encounter %>% group_by(eDWID,reviewyear,`Age Group`) %>% 
  summarise(cfrd = max(cfrd_cffpr),.groups = "drop") %>%
  group_by(reviewyear,`Age Group`) %>%
  summarise(Prevalence = mean(as.numeric(cfrd)-1,na.rm = T)*100,.groups = "drop") %>%
  pivot_wider(names_from = `Age Group`,values_from = Prevalence) %>%
  kable(.,digits = 3)
```

```{r}
#| label: tbl-prevalence_cffpr
#| tbl-cap: "Yearly Prevalence of CFRD Based on CFFPR Alone"
encounter %>% group_by(eDWID,reviewyear) %>% 
  summarise(cfrd = max(cfrd_cffpr),.groups = "drop") %>%
  group_by(reviewyear) %>%
  summarise(Prevalence = mean(as.numeric(cfrd)-1,na.rm = T)*100,.groups = "drop") %>%
  kable(.,digits = 3)
```

In @tbl-prevalence_cffpr_year and @tbl-prevalence_cffpr, prevalence is calculated as the total number of CFRD cases (by registry definition alone) seen in clinic divided by the total number of people seen in clinic in a given year.

```{r}
#| label: tbl-incidence_both
#| tbl-cap: "Year-on-Year Incidence of CFRD Based on CFFPR and OGTT"
# For each year, pull everyone with and without CFRD that year
year_inc = function(df=encounter,year,py = 1000){
  # Filter and count
  inc = df %>% 
    filter(reviewyear==year,is.na(first_cfrd_both) | first_cfrd_both == year) %>%
    group_by(eDWID) %>% filter(row_number() == 1) %>%
    select(eDWID,reviewyear,encounternum,`Age Group`,
           twohour_bloodglucose,cfrd_both,first_cfrd_both)
  inc = inc %>% group_by(cfrd_both) %>% count(`Age Group`,.drop = F)
  inc = inc[!is.na(inc$`Age Group`),]
  # Calculations
  new_cases = sum(inc$n[inc$cfrd_both=="Yes"])
  total_at_risk = sum(inc$n)
  total_inc = new_cases/(total_at_risk/py)

@@

```{r}
#| label: tbl-prevalence_both_year
#| tbl-cap: "Yearly Prevalence of CFRD by Age Group Based on CFFPR and OGTT"
encounter$cfrd_both = factor(encounter$cfrd_both,levels = c("No","Yes"),ordered = T)
encounter %>% group_by(eDWID,reviewyear,`Age Group`) %>% 
  summarise(cfrd = max(cfrd_both),.groups = "drop") %>%
  group_by(reviewyear,`Age Group`) %>%
  summarise(`Prevalence` = mean(as.numeric(cfrd)-1,na.rm = T)*100,.groups = "drop") %>%
  pivot_wider(names_from = `Age Group`,values_from = Prevalence)%>%
  kable(.,digits = 3)
```

```{r}
#| label: tbl-prevalence_both
#| tbl-cap: "Yearly Prevalence of CFRD Based on CFFPR and OGTT"
encounter %>% group_by(eDWID,reviewyear) %>% 
  summarise(cfrd = max(cfrd_both),.groups = "drop") %>%
  group_by(reviewyear) %>%
  summarise(`Prevalence` = mean(as.numeric(cfrd)-1,na.rm = T)*100,.groups = "drop") %>%
  kable(.,digits = 3)
```

In @tbl-prevalence_both_year and @tbl-prevalence_both, prevalence is calculated as the total number of CFRD cases (by either registry definition or OGTT) seen in clinic divided by the total number of people seen in clinic in a given year.

```{r}
#| label: tbl-screening_ogtt
#| tbl-cap: "Yearly OGTT Screening Rates"
year_screen = function(df=encounter,year){
  # Filter and count
  inc = df %>% 
    filter(reviewyear==year,first_cfrd_both >= year) %>%
    mutate(ogtt = !(is.na(ogttfast_bloodglucose) | is.na(twohour_bloodglucose))) %>%
    group_by(eDWID) %>%
    summarise(ogtt = max(ogtt),`Age Group` = `Age Group`[1])
  inc = inc %>% group_by(ogtt) %>% count(`Age Group`) %>% filter(`Age Group` != "< 10")
  # Calculations
  total_screened = sum(inc$n[inc$ogtt==1])
  total_seen = sum(inc$n)
  screen_rate = total_screened/total_seen * 100

@@

```{r}
#| label: tbl-screening_a1c
#| tbl-cap: "Yearly HbA1c Screening Rates"
year_screen = function(df=encounter,year){
  # Filter and count
  inc = df %>% 
    filter(reviewyear==year,first_cfrd_both >= year) %>%
    mutate(a1c = !is.na(hgba1c)) %>%
    group_by(eDWID) %>%
    summarise(a1c = max(a1c),`Age Group` = `Age Group`[1])
  inc = inc %>% group_by(a1c) %>% count(`Age Group`) %>% filter(`Age Group` != "< 10")
  # Calculations
  total_screened = sum(inc$n[inc$a1c==1])
  total_seen = sum(inc$n)
  screen_rate = total_screened/total_seen * 100

  screened_10_18 = inc$n[inc$a1c==1 & inc$`Age Group` == "10 - 18"]
  seen_10_18 = sum(inc$n[inc$`Age Group` == "10 - 18"])
  rate_10_18 = screened_10_18/seen_10_18 *100

  screened_19_30 = inc$n[inc$a1c==1 & inc$`Age Group` == "19 - 30"]
  seen_19_30 = sum(inc$n[inc$`Age Group` == "19 - 30"])
  rate_19_30 = screened_19_30/seen_19_30 *100

  screened_30 = inc$n[inc$a1c==1 & inc$`Age Group` == "> 30"]
  seen_30 = sum(inc$n[inc$`Age Group` == "> 30"])
  rate_30 = screened_30/seen_30 *100

@@

In @tbl-screening_ogtt and @tbl-screening_a1c, people with CFRD (by either registry definition or OGTT) and those under age 10 were excluded. Screening rate is calculated as the number of people with an OGTT divided by the total number of people seen in clinic that year.
